### 2022-11-22 draft.tex
Deploying to gh-pages from @ graciellehigino/ms_range_interactions@6a03212150abe2f94369e350bd8509868cfa9ea2 🚀
Deploying to gh-pages from @ graciellehigino/ms_range_interactions@1c8da06906e50354f2b96c92b09fa8cbaee445a9 🚀

It has been demonstrated that the agreement between range maps and point
data varies geographically (Hurlbert and Jetz 2007; Hurlbert and White
2005; Ficetola et al. 2014). Adding ecological interaction data to this
comparison might help to elucidate where these (dis)agreements are more
likely to be true and which dataset better represent the actual
distribution of species. In this context, we elaborate a method that
allows us to refine distribution data (more precisely range maps) based
on interaction data, considering the basic assumption that predators can
only be present in regions where they are connected to at least one
herbivore - and thus indirectly connected to primary producers. We used
a Serengeti food web dataset (Baskerville et al. 2011) (which comprises
carnivores, herbivores, and plants from Tanzania) to demonstrate how a
mismatch between occurrence and interaction data can highlight
significant uncertainty areas in IUCN range maps. Finally, we add the
GBIF occurrence points for the Serengeti species to the investigation,
discuss the mechanisms that can lead to the lack of agreement between
data, and build from that a vision for the next steps, reinforcing the
importance of geographically explicit interaction data.

@@

15540\strut

@@

0.560\strut

@@

0.670\strut

@@

43307\strut

@@

0.848\strut

@@

0.252\strut

@@

3873\strut

@@

0.916\strut

@@

11384\strut

@@

0.934\strut

@@

68137\strut

@@

7358\strut

@@

0.780\strut

@@

19872\strut

@@

0.995\strut

@@

47243\strut

@@

0.832\strut

@@

38856\strut

@@

0.979\strut

@@

5567\strut

@@

3695\strut

@@

26705\strut

@@

22380\strut

@@

11901\strut

@@

29963\strut

@@

0.999\strut

@@

17465\strut

@@

465\strut

@@

0.998\strut

@@

20852\strut

@@

0.991\strut

@@

36011\strut

@@

0.984\strut

@@

10579\strut

@@

20761\strut

@@

9650\strut

@@

7070\strut

@@

463\strut

@@

2303\strut

@@

17728\strut

@@

0.972\strut

@@

47697\strut

@@

5418\strut

@@

0.470\strut

@@

9654\strut

@@

4002\strut

@@

23171\strut

@@

0.938\strut

@@

25223\strut

@@

0.250\strut

@@

The jackal is a widespread taxon in northern Africa, Europe, and
Australasia, generally well adapted to local conditions due to its
largely varied diet (Tsunoda and Saito 2020; Krofel et al. 2021).
Because of that, we expected that the \emph{Canis} species in our
dataset would be the ones losing the least amount of range, with a
higher value of the proportion of GBIF pixels within their IUCN range
maps. However, the taxonomy of this group is a matter of intense
discussion, as molecular and morphological data seem to disagree in the
clustering of species and subspecies (Krofel et al. 2021; Stoyanov
2020). This debate probably influenced our results: with originally only
56.2\% of the GBIF pixels of the golden jackal (\emph{Canis aureus})
overlapping with the IUCN data, we suspect that many of the GBIF
occurrences refer to other \emph{Canis} species, and that its taxonomic
identification in the network database is probably outdated. This led to
a complete exclusion of \emph{Canis aureus} from its original range in
our analysis, despite the fact that this species has four documented
preys in our metaweb. This example illustrates how the taxonomic,
geographical and ecological data can be used to validate one another.

Here we show that when ecological interaction data (predator-prey
interactions within food webs) are used to refine species range maps,
there are significant reductions in the IUCN range size of predatory
organisms. Despite showing the potential importance of accounting for
species interactions when estimating the range of a species, it remains
unclear the extent to which the patterns observed represent ecological
processes or a lack of data.

\hypertarget{connectivity-diversity-and-range-preservation}{%
\subsection{Connectivity, diversity and range
preservation}\label{connectivity-diversity-and-range-preservation}}

In the Serengeti food web there is a positive relationship between the
out degrees of predators and the size of their ranges. Here, we showed
that there is a negative relationship between the relative loss of
predators' ranges and their number of preys, reinforcing the idea that
generalist species can preserve their distributions longer while losing
interactions. The factors limiting the geographical range of a species
in a community can vary with connectivity and richness (Svenning et al.
2014). Younger communities may be more affected by environmental
limitations because they are dominated by generalist species, while
older metacommunities are probably affected in different ways in the
center of the distribution, at the edge of ranges, and in sink and
source communities (Svenning et al. 2014; Godsoe et al. 2017; Cazelles
et al. 2016; Bullock et al. 2000). Additionally, it is likely that
species with larger ranges of distribution and those that are more
generalists would co-occur with a greater number of other species
(Dáttilo et al. 2020), while dispersal capacity of competitive species
modulate their aggregation in space and the effect of interactions on
their range limits (Godsoe et al. 2017).

@@

The geographical mismatch between predators and preys has ecological
consequences such as loss of ecosystem functioning and extinction of
populations (Anderson et al. 2016; Dáttilo and Rico-Gray 2018; Pringle
et al. 2016; Young et al. 2013). Climate change is one of the causes of
this, leading, for instance, to the decrease of plant populations due to
the lack of pollination (Bullock et al. 2000; Afkhami, McIntyre, and
Strauss 2014; Godsoe et al. 2017). However, this mismatch can also be
purely informational. When the distribution of predators and preys does
not superpose, it can mean we lack information about the distribution of
either species or about their interactions (e.g., predators may be
feeding on different species than the ones in our dataset outside the
Serengeti ecosystem). Here we addressed part of this problem by
comparing the IUCN range maps with GBIF occurrences, which helped us
clarify what is the shortfall for each species.

The lack of superposition between IUCN range maps and GBIF occurrences
suggests that we certainly do miss geographical information about the
distribution of a certain species, but this is not an indicator of the
completeness of the information about ecological interactions. However,
if both GBIF and IUCN occurrences tend to superpose and still the
species is locally removed, this indicates we don't have information
about all its interactions. The combination of this rationale with our
method of updating range maps based on ecological interactions allows us
to have a clearer idea of which information we are missing. For example,
the lion (\emph{Panthera leo}) was one of the species with no difference
between the original and the updated ranges, but 40.7\% of the GBIF
occurrences for this species fell outside its IUCN range
(fig.~\ref{fig:gbif}). In this particular case, the IUCN maps seem to
agree with species interaction data. However, the disagreement between
the IUCN and the GBIF databases is concerning and suggests that the IUCN
maps might underestimate the lion's distribution. On the other hand,
\emph{Leptailurus serval} and \emph{Canis mesomelas} are two of the
three species that lose the higher proportion of range due to the lack
of paths to a herbivore, but are also some of the species with the
higher proportion of GBIF occurrences inside IUCN range maps
(fig.~\ref{fig:gbif}). This indicates that the information we are
missing for these two species is related to either the occurrence of an
interaction or the presence of interacting species. To illustrate that,
we mapped the GBIF data for the prey of \emph{Leptailurus serval}, with
a mobility buffer around each point (fig.~\ref{fig:serval}). When
considering GBIF data, approximately 42\% of the prey's occurrences are
within the portion of the serval's range that was lost. With the buffer
area, this corresponds to 15\% of the lost range. This means that by
adding GBIF information, we would reduce the loss of range (or
information) for the predator by 15\% since its distribution is
conditional on the occurrence of its preys.

@@

Finally, the extreme case of \emph{Canis aureus} illustrates a lack of
both geographical and ecological information: only half of its GBIF
presence pixels and none of its preys occur inside its IUCN range. We
believe, therefore, that the validation of species distribution based on
ecological interaction is a relevant method that can further fill in
information gaps. Nevertheless, it is imperative that more
geographically explicit data about ecological networks and interactions
become available. This would help clarify when cooccurrences can be
translated into interactions (Windsor et al. 2022) and help the
development of more advanced validation methods for occurrence data.

@@

\leavevmode\hypertarget{ref-Anderson2016SpaDis}{}%
Anderson, T. Michael, Staci White, Bryant Davis, Rob Erhardt, Meredith
Palmer, Alexandra Swanson, Margaret Kosmala, and Craig Packer. 2016.
{``The Spatial Distribution of African Savannah Herbivores: Species
Associations and Habitat Occupancy in a Landscape Context.''}
\emph{Philosophical Transactions of the Royal Society B: Biological
Sciences} 371 (1703): 20150314.
\url{https://doi.org/10.1098/rstb.2015.0314}.

@@

\leavevmode\hypertarget{ref-Bullock2000GeoSep}{}%
Bullock, James M., Rebecca J. Edwards, Peter D. Carey, and Rob J. Rose.
2000. {``Geographical Separation of Two Ulex Species at Three Spatial
Scales: Does Competition Limit Species' Ranges?''} \emph{Ecography} 23
(2): 257--71. \url{https://doi.org/10.1111/j.1600-0587.2000.tb00281.x}.

@@

\leavevmode\hypertarget{ref-Cazelles2016IntBio}{}%
Cazelles, Kévin, Nicolas Mouquet, David Mouillot, and Dominique Gravel.
2016. {``On the Integration of Biotic Interaction and Environmental
Constraints at the Biogeographical Scale.''} \emph{Ecography} 39 (10):
921--31. \url{https://doi.org/10.1111/ecog.01714}.

@@

\leavevmode\hypertarget{ref-Dattilo2020SpeDri}{}%
Dáttilo, Wesley, Nathalia Barrozo-Chávez, Andrés Lira-Noriega, Roger
Guevara, Fabricio Villalobos, Diego Santiago-Alarcon, Frederico Siqueira
Neves, Thiago Izzo, and Sérvio Pontes Ribeiro. 2020. {``Species-Level
Drivers of Mammalian Ectoparasite Faunas.''} \emph{Journal of Animal
Ecology} 89 (8): 1754--65.
\url{https://doi.org/10.1111/1365-2656.13216}.

\leavevmode\hypertarget{ref-Dattilo2018EcoNet}{}%
Dáttilo, Wesley, and Victor Rico-Gray, eds. 2018. \emph{Ecological
Networks in the Tropics: An Integrative Overview of Species Interactions
from Some of the Most Species-Rich Habitats on Earth}. 1st ed. 2018.
Cham: Springer International Publishing : Imprint: Springer.
\url{https://doi.org/10.1007/978-3-319-68228-0}.

@@

\leavevmode\hypertarget{ref-Pringle2016LarHer}{}%
Pringle, Robert M., Kirsten M. Prior, Todd M. Palmer, Truman P. Young,
and Jacob R. Goheen. 2016. {``Large Herbivores Promote Habitat
Specialization and Beta Diversity of African Savanna Trees.''}
\emph{Ecology} 97 (10): 2640--57.
\url{https://doi.org/10.1002/ecy.1522}.

@@

\leavevmode\hypertarget{ref-Svenning2014InfInt}{}%
Svenning, Jens Christian, Dominique Gravel, Robert D. Holt, Frank M.
Schurr, Wilfried Thuiller, Tamara Münkemüller, Katja H. Schiffers, et
al. 2014. {``The Influence of Interspecific Interactions on Species
Range Expansion Rates.''} \emph{Ecography} 37 (12): 1198--1209.
\url{https://doi.org/10.1111/j.1600-0587.2013.00574.x}.

@@

\leavevmode\hypertarget{ref-Young2013EffMam}{}%
Young, Hillary S., Douglas J. McCauley, Kristofer M. Helgen, Jacob R.
Goheen, Erik Otárola-Castillo, Todd M. Palmer, Robert M. Pringle, Truman
P. Young, and Rodolfo Dirzo. 2013. {``Effects of Mammalian Herbivore
Declines on Plant Communities: Observations and Experiments in an
\textsc{A} Frican Savanna.''} Edited by Luis Santamaria. \emph{Journal
of Ecology} 101 (4): 1030--41.
\url{https://doi.org/10.1111/1365-2745.12096}.

>>>>

It has been demonstrated that the agreement between range maps and point
data varies geographically (Hurlbert and Jetz 2007; Hurlbert and White
2005; Ficetola et al. 2014). Adding ecological interaction data to this
comparison might help to elucidate where these (dis)agreements are more
likely to be true and which dataset better represent the actual
distribution of species. In this context, we elaborate a method that
allows us to detect areas of potential misestimation of species'
distribution data (more precisely range maps) based on interaction data,
considering the basic assumption that predators can only be present in
regions where they are connected to at least one herbivore - and thus
indirectly connected to primary producers. We used a Serengeti food web
dataset (Baskerville et al. 2011) (which comprises carnivores,
herbivores, and plants from Tanzania) to demonstrate how a mismatch
between occurrence and interaction data can highlight significant
uncertainty areas in IUCN range maps. Finally, we add the GBIF
occurrence points for the Serengeti species to the investigation,
discuss the mechanisms that can lead to the lack of agreement between
data, and build from that a vision for the next steps, reinforcing the
importance of geographically explicit interaction data.

@@

9250\strut

@@

0.437\strut

@@

0.618\strut

@@

4822\strut

@@

0.844\strut

@@

0.253\strut

@@

427\strut

@@

0.918\strut

@@

1274\strut

@@

0.935\strut

@@

7563\strut

@@

816\strut

@@

0.782\strut

@@

2201\strut

@@

0.994\strut

@@

5239\strut

@@

0.833\strut

@@

4319\strut

@@

0.978\strut

@@

626\strut

@@

419\strut

@@

2961\strut

@@

2484\strut

@@

1318\strut

@@

3331\strut

@@

@@

1935\strut

@@

53\strut

@@

@@

2316\strut

@@

0.990\strut

@@

3999\strut

@@

0.985\strut

@@

1167\strut

@@

2307\strut

@@

1074\strut

@@

786\strut

@@

51\strut

@@

261\strut

@@

1961\strut

@@

0.973\strut

@@

5312\strut

@@

607\strut

@@

0.473\strut

@@

1078\strut

@@

443\strut

@@

2571\strut

@@

0.937\strut

@@

2808\strut

@@

0.251\strut

@@

Here we identify areas of data mismatch between species range maps by
using ecological interaction data (predator-prey interactions within
food webs). Our results did show a significant mismatch in the IUCN
range areas of specialized and generalist predatory organisms and their
prey, which highlights the importance of accounting for species
interactions when estimating the range of a species. Although this type
of data mismatch can result of actual ecological processes, outdated
occurrence data, taxonomic errors and more, we argue that, here, they
rather indicate a lack of interaction sampling data.

The case of the golden jackal (\emph{Canis aureus}) is a good
illustration of how the taxonomic, geographical and ecological data can
be used to validate one another. The jackal is a widespread taxon in
northern Africa, Europe, and Australasia, generally well adapted to
local conditions due to its largely varied diet (Tsunoda and Saito 2020;
Krofel et al. 2021). Because of that, we expected that
the~\emph{Canis}~species in our dataset would be the ones losing the
least amount of range, with a higher value of the proportion of GBIF
pixels within their IUCN range maps. However, the taxonomy of this group
is a matter of intense discussion, as molecular and morphological data
seem to disagree in the clustering of species and subspecies (Krofel et
al. 2021; Stoyanov 2020). This debate probably influenced our results:
with originally only 56.2\% of the GBIF pixels of the golden jackal
overlapping with its IUCN data, we suspect that many of the GBIF
occurrences refer to other~\emph{Canis}~species, and that its taxonomic
identification in the network database is probably outdated. This led to
a complete exclusion of~\emph{Canis aureus}~from its original range in
our analysis, despite the fact that this species has four documented
preys in our metaweb.

@@

The lack of superposition between IUCN range maps and GBIF occurrences
in our results suggests that we certainly miss geographical information
about the distribution of either the prey or the predator. On the other
hand, if both GBIF and IUCN occurrences tended to superpose and the
species was still locally removed, this indicates that we don't have
information about all its interactions (e.g., predators may be feeding
on different species than the ones in our dataset outside the Serengeti
ecosystem). This rationale can be illustrated with three types of
mismatches identified in our results.

First, \emph{Panthera leo} was one of the species with no difference
between ranges before and after our analysis, but 40.7\% of the GBIF
occurrences for this species fell outside its IUCN range
(fig.~\ref{fig:gbif}). In this particular case, the IUCN maps seem to
agree with species interaction data. However, the disagreement between
the IUCN and the GBIF databases is concerning and suggests that the IUCN
maps might underestimate the lion's distribution.

On the other hand,~\emph{Leptailurus serval}~and~\emph{Canis
mesomelas}~are two of the three species that have the higher proportion
of mismatched range due to the lack of paths to a herbivore, but are
also some of the species with the higher proportion of GBIF occurrences
inside IUCN range maps (fig.~\ref{fig:gbif}). This indicates that the
information we are missing for these two species is related to either an
additional interaction or to the presence of interacting species. To
illustrate that, we mapped the GBIF data for the prey
of~\emph{Leptailurus serval}, with a mobility buffer around each point
(fig.~\ref{fig:serval}). When considering GBIF data, approximately 42\%
of the prey's occurrences are within the portion of the predator's range
that was divergent from its original IUCN data. With the buffer area,
this corresponds to 15\% of the mismatched area. By adding GBIF
information for the prey, we could therefore reduce the discrepancy of
the range (or information) for the predator by 15\% since its
distribution is conditional on the occurrence of its preys. In other
words, the range mismatch was exagerated because we were missing
information on the presence of an interacting species (i.e., this also
indicates that there is a mismatch - or complementarity - between the
IUCN and GBIF data for their prey).

@@

Finally, the extreme case of~\emph{Canis aureus}~illustrates a lack of
both geographical and ecological information: only half of its GBIF
presence pixels and none of its preys occur inside its IUCN range. We
believe, therefore, that the validation of species distribution based on
ecological interaction is a relevant method that can further fill in
information gaps. Nevertheless, it is imperative that more
geographically explicit data about ecological networks and interactions
become available. This would help clarify when cooccurrences can be
translated into interactions (Windsor et al. 2022) and help the
development of more advanced validation methods for occurrence data.

@@

@@

@@

@@

@@

@@

@@
